---
title: "Statistical Analysis"
author: "Brandon Bowers"
date: "2025-03-19"
output: html_document
---
```{r}
 library(nlme)
 library(car)
```


```{r}
WD = "C:\\Users\\Bvipe\\OneDrive - Walla Walla University\\Desktop\\Bowers_2025_Octopus_Vestibular"

datelist_u = c("7_24","7_28","7_30","8_1","8_3","8_5","8_7","8_9","8_11","8_13","8_15","8_17")
```

#Read in Trial-type Log
```{r}
Trial_type_log = read.csv("C:\\Users\\Bvipe\\OneDrive - Walla Walla University\\Desktop\\Bowers_2025_Octopus_Vestibular\\Raw Data\\Trial Type.csv")
```

```{r}
FV_get_file = function(WD,d){ 
  wd = c(WD,"Processed Data","Full Video Results",d)
  wd_path = stringr::str_c(wd, collapse = "/")
  in_files_list = list.files(wd_path, pattern = NULL, all.files = FALSE,
           full.names = FALSE, recursive = FALSE,
           ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)
  return(in_files_list)
}
```

```{r}
PA_get_file = function(WD){ 
  wd = c(WD,"Processed Data\\Path Analysis Results")
  wd_path = stringr::str_c(wd, collapse = "/")
  in_files_list = list.files(wd_path, pattern = NULL, all.files = FALSE,
           full.names = FALSE, recursive = FALSE,
           ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)
  return(in_files_list)
}
```

#Create a table from the the processed data that makes all effects accessible. FULL VID RESULTS
```{r}
Date = c()
OC_ID = c()
Trial_Type = c()
Avg_Speed_Per_Bodylength = c()
Avg_Bearing_Change = c()
Avg_Bearing_vs_Heading_FV = c()

for (d in datelist_u){
  FV_files = FV_get_file(WD,d)
  for (f in FV_files){
    Date = c(Date,d)
    
    oc = unlist(strsplit(f,".csv"))[1]
    oc = unlist(strsplit(oc,paste0("O")))[2]
    OC_ID = c(OC_ID,oc)
    
    curr_data = read.csv(paste0(WD,"\\Processed Data\\Full Video Results\\",d,"\\",f))
    
    #Find average speed/bodylength and append it to master list
    avg_SPB = sum(curr_data$Speed_per_Bodylength) / length(curr_data$Speed_per_Bodylength)
    Avg_Speed_Per_Bodylength = c(Avg_Speed_Per_Bodylength,avg_SPB)
    
    #Find average bearing change
    avg_BC = sum(curr_data$Bearing_Change) / length(curr_data$Bearing_Change)
    Avg_Bearing_Change = c(Avg_Bearing_Change,avg_BC)
    
    #Find avg bearing vs heading
    avg_BvH = sum(curr_data$Bearing_Movedir_Dif) / length(curr_data$Bearing_Movedir_Dif)
    Avg_Bearing_vs_Heading_FV = c(Avg_Bearing_vs_Heading_FV,avg_BvH)
    
    for (t in 1:length(Trial_type_log$Trial.Type)){
      if (Trial_type_log$Date[t] == d & as.character(Trial_type_log$OC_ID[t]) == oc){
        Trial_Type = c(Trial_Type,Trial_type_log$Trial.Type[t])
      }
    }
  }
}

FV_master_set = data.frame(Date,OC_ID,Trial_Type, Avg_Speed_Per_Bodylength,Avg_Bearing_Change,Avg_Bearing_vs_Heading_FV)

```

#Create a table from the the processed data that makes all effects accessible. PATH ANALYSIS RESULTS
```{r}
PA_files = PA_get_file(WD)
Date = c()
OC_ID = c()
Trial_Type = c()
Movement_Type = c()
Distance_of_Movement = c()
Duration = c()
Speed = c()
Speed_per_Bodylength = c()
Bearing_Change = c()
Avg_Bearing_vs_Heading = c()
Avg_Bearing_vs_Heading_Change = c()

for (fi in PA_files){
  current_data = read.csv(paste0(WD,"\\Processed Data\\Path Analysis Results\\",fi))
  Date = c(Date,current_data$date)
  OC_ID = c(OC_ID,current_data$oc_ID)
  Movement_Type = c(Movement_Type, current_data$movement_type)
  Distance_of_Movement = c(Distance_of_Movement ,current_data$total_movement_dis)
  Duration = c(Duration, current_data$duration)
  Speed = c(Speed,current_data$speed)
  Speed_per_Bodylength = c(Speed_per_Bodylength, current_data$speed_per_bodylength)
  Bearing_Change = c(Bearing_Change, current_data$angle_dev)
  Avg_Bearing_vs_Heading = c(Avg_Bearing_vs_Heading, current_data$avg_bear_head_dif)
  Avg_Bearing_vs_Heading_Change = c(Avg_Bearing_vs_Heading_Change, current_data$avg_bear_head_dif_dif)
  
  count = 0
  for (t in 1:length(Trial_type_log$Trial.Type)){
    if (Trial_type_log$Date[t] == current_data$date[1]){
      for (l in 1:length(current_data$oc_ID)){
        if (current_data$oc_ID[l] == Trial_type_log$OC_ID[t]){
          Trial_Type = c(Trial_Type,Trial_type_log$Trial.Type[t])
          count = count + 1
        }
      }
    }
  }
}

PA_master_set = data.frame(Date,OC_ID,Trial_Type,Movement_Type,Distance_of_Movement,Duration,Speed,Speed_per_Bodylength,Bearing_Change,Avg_Bearing_vs_Heading,Avg_Bearing_vs_Heading_Change)
```

########
#Run mixed effects model#
########

#Make the relationship between relevant variables linear by taking their logs.
```{r}

```

#Set orthagonal contrasts
```{r}

```

#Use a Type III sum of squares
```{r}
PA_master_set$Trial_Type=as.factor(PA_master_set$Trial_Type)

bearing_change.lme=lme(abs(Bearing_Change)~Trial_Type,random=~1|OC_ID,
                       data=PA_master_set[PA_master_set$Movement_Type==1,])
bearing_change.ANOVA=Anova(bearing_change.lme,type="III")
bearing_change.ANOVA
```
```{r}
PA_master_set$Trial_Type=as.factor(PA_master_set$Trial_Type)

bearing_change.lme=lme(abs(Speed)~Trial_Type,random=~1|OC_ID,
                       data=PA_master_set[PA_master_set$Movement_Type==1,])
bearing_change.ANOVA=Anova(bearing_change.lme,type="III")
bearing_change.ANOVA
```
```{r}
PA_master_set$Trial_Type=as.factor(PA_master_set$Trial_Type)

bearing_change.lme=lme(abs(Avg_Bearing_vs_Heading)~Trial_Type,random=~1|OC_ID,
                       data=PA_master_set[PA_master_set$Movement_Type==1,])
bearing_change.ANOVA=Anova(bearing_change.lme,type="III")
bearing_change.ANOVA
```
```{r}
PA_master_set$Trial_Type=as.factor(PA_master_set$Trial_Type)

bearing_change.lme=lme(abs(Avg_Bearing_vs_Heading_Change)~Trial_Type,random=~1|OC_ID,
                       data=PA_master_set[PA_master_set$Movement_Type==1,])
bearing_change.ANOVA=Anova(bearing_change.lme,type="III")
bearing_change.ANOVA
```

```{r}
FV_master_set$Trial_Type=as.factor(FV_master_set$Trial_Type)

bearing_change.lme=lme(Avg_Speed_Per_Bodylength~Trial_Type,random=~1|OC_ID, 
                       correlation=corAR1(form= ~Date|OC_ID),
                       data=FV_master_set)
bearing_change.ANOVA=Anova(bearing_change.lme,type="III")
bearing_change.ANOVA
```

